---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## The data was downloaded from the web (url:"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") and stored in the "PeerAssessment 1" folder. 

## Loading and preprocessing the data
```{r, echo = TRUE}
setwd("~/Desktop/Data_Science_Course/Reproducible Research/PeerAssessment 1")
unzip("repdata-data-activity.zip", exdir =".", unzip = "internal", setTimes = TRUE)
myfile <- read.csv("activity.csv")
library(dplyr) 
library(ggplot2)
mytable <- myfile %>% group_by(date) %>% summarise(steps= sum(steps))
p1 <- ggplot(data = mytable) + 
+ aes(x = date, y = steps)+
+ geom_histogram(stat = "identity", fill = "black")
 p1 + theme(axis.text.x = element_text(angle = 45))
 summary(mytable)
 ```
 ## The mean steps per day is 10766 and median is 10765. 
 
 ## What is the average daily activity pattern?
 
 ## Average steps taken per day 
 mytable2 <- myfile %>% group_by(interval, date) %>% summarise(steps = mean(steps))
 plot(mytable2$interval, mytable2$steps, xlab = "interval", ylab = "mean steps per day",  type = "l")
 summary(mytable2)
 
 ##  as in the summary of mytable2,we get the maximum number of steps taken in an interval is 806. It is confirmed by 
 ```{r, echo = TRUE}
 max_steps <- max(mytable2$steps, na.rm = TRUE)
 max_steps
 ```
 
 ## The corresponding 5-minute interval is 615, as detected by :
 ```{r. echo= TRUE}
 mytable2$interval[match(806, mytable2$steps)]
 ```
 ## Imputing missing values
For simplicity, the missing values for the steps are replaced with the mean steps/interval as follows
 ```{r, echo = TRUE}
 newtable <- mytable ## makes a copy the original data
 replacement <- mean(mytable$steps, na.rm = TRUE) ## calculates the mean for steps to be used as replacement
 newtable$steps[is.na(newtable$steps)] <- replacement ## replaces the missing values with the replacement value
 miss_val <- sum(is.na(newtable$steps))  ## to confirm there is no missing value
 miss_val

 p2 <- ggplot(data = newtable) +             ## plotting the histogram
 aes(x = date, y = steps)+
 geom_histogram(stat = "identity", fill = "black")
 p2 + theme(axis.text.x = element_text(angle = 45))
```
 ## The mean and median steps are 
 ```{r, echo= TRUE}
 newmeansteps <- mean(newtable$steps)
 newmediansteps <- median(newtable$steps)
 print(c(newmeansteps,newmediansteps))
  ```
  
 ## to compare the mean, median and other features in the two tables
 ```{r, echo = TRUE}
 summary(mytable)
 summary(newtable) 
 ```
 ## The mean steps per day is 10766 and median is 10766. 
 ## There is no change in the mean. The median changes by 1 step only. 
 ## There is no significant impact for imputing the missing data on the estimates of the total daily number of steps.
 
 
 Are there differences in activity patterns between weekdays and weekends?
 ```{r, echo = TRUE}
 levelstable <- newtable
 
 
# Are there differences in activity patterns between weekdays and weekends?
