---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## The data was downloaded from the web (url:"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") and stored in the "PeerAssessment 1" folder. 

## Loading and preprocessing the data
```{r, echo = TRUE}
setwd("~/Desktop/Data_Science_Course/Reproducible Research/PeerAssessment 1")
unzip("repdata-data-activity.zip", exdir =".", unzip = "internal", setTimes = TRUE)
myfile <- read.csv("activity.csv")
library(dplyr) 
library(ggplot2)
mytable <- myfile %>% group_by(date) %>% summarise(steps= sum(steps))
p1 <- ggplot(data = mytable) + 
+ aes(x = date, y = steps)+
+ geom_histogram(stat = "identity", fill = "black")
 p1 + theme(axis.text.x = element_text(angle = 45))
```



## What is mean total number of steps taken per day?
```{r, echo = TRUE}
 summary(mytable)
 ```
 ## The mean steps per day is 10766 and the median is 10765. 


## What is the average daily activity pattern?
```{r, echo = TRUE}
mytable2 <- myfile %>% group_by(interval, date) %>% summarise(steps = mean(steps))
plot(mytable2$interval, mytable2$steps, xlab = "interval", ylab = "mean steps per day",  type = "l")
summary(mytable2)
```
## The time series plot looks different than the example because of the interval setting as in the assignment. 
## From the summary of mytable2, we get the maximum number of steps taken in an interval is 806.
## The corresponding 5-minute interval for the maximum number of steps taken is 615, as detected by :
 ```{r. echo= TRUE}
 mytable2$interval[match(806, mytable2$steps)]
 ```
## The total number of missing vlaues(NA's) is 2304 (from the summary of mytable2)


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
